---
title: "harrington_statcomp_hw5"
format: 
  html: 
    output-file: "harrington_statcomp_hw5.html"
---

```{r}
#| echo: false
#| warning: false

library(dplyr)
library(tidyr)
library(ggplot2)
library(reactable)
library(plotly)
library(crosstalk)
```

```{r}
#| code-fold: true
#| warning: false

df <- data.frame(
x = c(5, 1, 4, NA, NA, NA, 2, 1),
y = c(5, NA, 3, 2, 5, NA, 2, 1),
z = c(1, 4, 2, 5, 2, NA, 5, 4),
t = c(4, 2, NA, 1, 5, 2, 3, 3)
)

cor(df$y, df$x, use = "pairwise.complete.obs")
cor(df$y, df$z, use = "pairwise.complete.obs")
cor(df$y, df$t, use = "pairwise.complete.obs")

meanu1= mean(df[,1], na.rm= T)
meanu2 = mean(df[,2], na.rm= T)
meanu3 = mean(df[,3], na.rm = T)
meanu4= mean(df[,4], na.rm= T)

u14 = meanu2 - meanu4 + df[6,4]
u14

```

```{r}
#| code-fold: true
#| warning: false

df2 <- rbind(
x = c(5, 1, 4, NA, NA, NA, 2, 1),
y = c(5, NA, 3, 2, 5, 2.142857, 2, 1),
z = c(1, 4, 2, 5, 2, NA, 5, 4),
t = c(4, 2, NA, 1, 5, 2, 3, 3)
)
df2 = as.data.frame(df2)

movie_means = c(0,0,0,0,0,0,0,0)
for (i in 1:8) {
  movie_means[i] = mean(df2[,i], na.rm = T)
}
movie_means
```
The above is a vector containing the means of the movie reviews.
```{r}
#| code-fold: true
#| warning: false
movie_sds = c(0,0,0,0,0,0,0,0)
for (i in 1:8) {
  movie_sds[i] = sd(df2[,i], na.rm = T)
}
movie_sds
```
The above is a vector containing the standard deviations of the movie reviews.
```{r}
#| code-fold: true
#| warning: false
for (i in 1:8) {
df2[,i] = (df2[,i]- movie_means[i])/movie_sds[i]
}
df2
```
The standardized reviews of the movies.
```{r}
#| code-fold: true
#| warning: false

cor(df2,use = "pairwise.complete.obs", method = 'pearson')

```
Note that in the above matrix, two different movies with a correlation of 1 occurs because they share only 2 reviewers, and thus are guaranteed to have a linear relation. These entries should not be considered. Thus, the most similar movies are 1 and 5, 7 and 8, and 2 and 8.


```{r}
#| code-fold: true
#| warning: false

data= read.csv("http://manctsui.github.io/datasets/countries.csv")

data2 <- data |>
 group_by(Region) |>
 mutate(
 Agriculture = ifelse(is.na(Agriculture),
 mean(Agriculture, na.rm = TRUE), Agriculture)
 ) |>
    mutate(
 Industry = ifelse(is.na(Industry),
 mean(Industry, na.rm = TRUE), Industry)) |>
 ungroup()

data2 = data2 |> 
  select(Country, Region, Industry, Agriculture)

reactable(data2, defaultPageSize = 6, searchable = T)

sum(is.na(data$Agriculture))
sum(is.na(data$Industry))
```
# The original dataset had 15 missing values for Agriculture, and 16 missing values for Industry

```{r}
#| code-fold: true
#| warning: false


df3 <- data.frame(
temp = c(79, 74, 73, 76, 82, 72, 71, 79, 83, 70, 73, 75, 77),
chirps = c(44, 40, 38, 43, 48, 37, 34, 45, 51, 32, 35, 42, 43)
)
lm(temp ~ chirps, data = df3)


```

```{r}
rho = cor(df3$temp, df3$chirps)
mu1 = mean(df3$temp)
sd1 = sd(df3$temp) 
mu2 = mean(df3$chirps)  
sd2 = sd(df3$chirps)

beta1 = rho*(sd1/sd2)

alpha = mu1 - beta1*mu2
```
By multiplying the correlation of temp and chirps by the standard deviation of temp, and then dividing by the standard deviation for chirps, we get beta, the coefficient for the line of best fit. Then, by multiplying this beta with the mean of chirps and subtracting that from the mean of temp, we get alpha, the y-intercept for the line of best fit.