---
title: "Harrington Stat Comp HW 3"
format: 
  html: 
    output-file: "hw3.html"
---

```{r}
#| echo: false
#| warning: false

library(dplyr)
library(tidyr)
library(reactable)


```

# 1) The following is a table displaying information on various Universities around the world.

```{r}
#| code-fold: true
#| warning: false
#| 
df= read.csv('http://manctsui.github.io/datasets/timesData.csv')
reactable(df, defaultPageSize = 4, searchable = T)
```

# 2) Dr. Tsui's alma mater, UCLA, had the following world rankings.

```{r}
#| code-fold: true
#| warning: false
#| 
df |>
  filter(grepl("Los Angeles", university_name) ==T) |>
  select(university_name ,world_rank, year)

```

# 3-4) UCLA vs. UC Berkeley Rankings

```{r}
#| code-fold: true
#| warning: false
#| 
#| layout-ncol: 2
#| 
x= df |> 
  filter(grepl("Los Angeles", university_name) ==T) |>
  select(year)
y= df |> 
  filter(grepl("Los Angeles", university_name) ==T) |>
  select(world_rank)
x_plot= x[,1]
y_mid= y[,1]
y_plot=as.numeric(y_mid)
plot(x_plot,y_plot,
     main = "Ranking of UCLA",
     xlab= "Year",
     ylab = "Rank",
     type= "l",
     xlim = c(2011,2016),
     ylim = c(5,16))

x= df |> 
  filter(grepl("Berkeley", university_name) ==T) |>
  select(year)
y= df |> 
  filter(grepl("Berkeley", university_name) ==T) |>
  select(world_rank)
x_plot= x[,1]
y_mid= y[,1]
y_plot=as.numeric(y_mid)
plot(x_plot,y_plot,
     main = "Ranking of UC Berkeley",
     xlab= "Year",
     ylab = "Rank",
     type= "l",
     xlim = c(2011,2016),
     ylim = c(5,16))
```

# 5) The following are the US Universities ranked in the top 50 globally, grouped by year

```{r}
#| code-fold: true
#| warning: false
#| 
#| layout-ncol: 6
#| 
q5 = df |> 
  filter(country == "United States of America", as.numeric(world_rank)<= 50, year == "2011") |>
  select(university_name, world_rank, year)

reactable(q5, defaultPageSize = 50, searchable = T)

q5 = df |> 
  filter(country == "United States of America", as.numeric(world_rank)<= 50, year == "2012") |>
  select(university_name, world_rank, year)

reactable(q5, defaultPageSize = 50, searchable = T)

q5 = df |> 
  filter(country == "United States of America", as.numeric(world_rank)<= 50, year == "2013") |>
  select(university_name, world_rank, year)

reactable(q5, defaultPageSize = 50, searchable = T)

q5 = df |> 
  filter(country == "United States of America", as.numeric(world_rank)<= 50, year == "2014") |>
  select(university_name, world_rank, year)

reactable(q5, defaultPageSize = 50, searchable = T)

q5 = df |> 
  filter(country == "United States of America", as.numeric(world_rank)<= 50, year == "2015") |>
  select(university_name, world_rank, year)

reactable(q5, defaultPageSize = 50, searchable = T)

q5 = df |> 
  filter(country == "United States of America", as.numeric(world_rank)<= 50, year == "2016") |>
  select(university_name, world_rank, year)

reactable(q5, defaultPageSize = 50, searchable = T)
```
We can see that many of the Universities ranked above 30th in 2011 do not break the top 50 in 2016.




# The following are additional tables to review the top 50 US Universities in other ways

```{r}
#| code-fold: true
#| warning: false
q5 = df |> 
    group_by(year, world_rank) |>
  filter(country == "United States of America", as.numeric(world_rank)<= 50) |>
  select(university_name, world_rank, year)

reactable(q5, defaultPageSize = 10, searchable = T)
```

```{r}
#| code-fold: true
#| warning: false
q6 = df |> 
    group_by(year, world_rank) |>
  filter(as.numeric(world_rank)<= 50) |>
  select(university_name, country, world_rank, year)

reactable(q6, defaultPageSize = 50, searchable = T)
```
